---
title: "Titanic Project"
author: "Stavros Oikonomou"
date: "22/3/2020"
output: html_document
---

### Loading the libraries
```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(dplyr)
library(randomForest)
library(caret)
```

### Loading the data
```{r}
setwd("C:/Users/tit0v/R/titanic-kaggle-project")

titanic.train <- read.table("train.csv", sep = ','
                                  ,stringsAsFactors = TRUE, header = TRUE,na.strings =c(""," "))
titanic.test <- read.table("test.csv", sep = ','
                            ,stringsAsFactors = TRUE, header = TRUE,na.strings =c(""," "))
```

### Creating a extra column to help me split the dataset after bind
```{r}
titanic.train$istrain <- T
titanic.test$istrain <- F
```
### Adding the survived column for test set
```{r}
titanic.test$Survived <- NA
```
### binding the train and test set
```{r}
titanic.full <- rbind(titanic.train,titanic.test)
```
### Categorical casting
```{r}
str(titanic.full)
titanic.full$Pclass <- as.factor(titanic.full$Pclass)
titanic.full$Parch <- as.factor(titanic.full$Parch)
titanic.full$SibSp <- as.factor(titanic.full$SibSp)
```
### removing cabin column
```{r}
titanic.full$Cabin <- NULL
```
### Number of NA
```{r}
sapply(titanic.full, function(x) sum(is.na(x)))
na.values.df <-sapply(titanic.full, function(x) sum(is.na(x)))
na.values.df[na.values.df>0]

```
### removing Embarked NA
```{r}
titanic.full[is.na(titanic.full$Embarked), "Embarked"] <- "S"
levels(titanic.train$Embarked)
table(titanic.full$Embarked)
```
### Removing NA from Fare
```{r}
median(titanic.full$Fare, na.rm = T)
titanic.full[is.na(titanic.full$Fare), ]
titanic.full$Fare[is.na(titanic.full$Fare)] <- median(titanic.full$Fare, na.rm = T)
```
### removing NA from Age with Linear Regression
```{r}
top_quant <- boxplot.stats(titanic.full$Age)$stats[5]
agee <- titanic.full[titanic.full$Age <= top_quant,]
na.age <- titanic.full[is.na(titanic.full$Age),]
lm_model <- lm(Age~Sex + SibSp + Pclass,
               data=agee)

age_pred <- predict(lm_model, newdata =na.age[c("Sex", "SibSp","Pclass")]  )
titanic.full[is.na(titanic.full$Age),]$Age <- round(age_pred)
```
### Extracting the Title from Names
```{r}
titanic.full$Title <- vector("character",length=nrow(titanic.full))
for (i in 1:nrow(titanic.full)) {
   x <- as.character(titanic.full$Name[i])
   m <- regexec(",(\\s+\\w+)+\\.", x) 
   titanic.full$Title[i] <- unlist(strsplit(unlist(regmatches(x,m))," "))[2]
}
levels(as.factor(titanic.full$Title))
titanic.full$Title[titanic.full$Title %in% c("Don.","Rev.","Dr.","Major.","Master.", "Sir.","Col.","Capt.","Jonkheer.")] <- "High class male"
titanic.full$Title[titanic.full$Title %in% c("Dona.","Lady.","Mlle.","Mme.", "the", "Ms.")] <- "high class female"
```
### Split the data and loading the y_test
```{r}
titanic.train <- titanic.full[titanic.full$istrain==T,]
titanic.test <- titanic.full[titanic.full$istrain==F,]
titanic.train$Survived <- as.factor(titanic.train$Survived)
titanic.train$Title <- as.factor(titanic.train$Title)
titanic.test$Title <- as.factor(titanic.test$Title)
str(titanic.train)

y_test<- read.table("gender_submission.csv", sep = ','
                    ,stringsAsFactors = TRUE, header = TRUE)
```

### Random Forest
#### Fitting the model, as we can see there is no difference in OOB error after 100 trees
```{r}
rf_model <- randomForest(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title,
                         data=titanic.train,nodesize = 0.01 * nrow(titanic.test))
plot(rf_model)
```


```{r}
tcontrol <- trainControl(method = "cv", number = 10, repeats = 3)
tunegrid <- expand.grid(mtry = seq(2,12,2))
modelRF <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title, ntree=100,
                 data=titanic.train, method = "rf",trControl = tcontrol, tuneGrid=tunegrid)
print(modelRF)
plot(modelRF)
```

```{r}
plot(modelRF)
```

#### So we gonna choose 100 trees and 10 features

### Random Forest
```{r}
rf2_model <- randomForest(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title, ntree=100,
                         mtry=10,data=titanic.train,nodesize = 0.01 * nrow(titanic.test))
pred_test <- predict(rf2_model, 
                   newdata = titanic.test[c("Pclass","Sex","Age","SibSp","Parch","Fare","Embarked","Title")])
pred_train <- predict(rf_model, 
                     newdata = titanic.train[c("Pclass","Sex","Age","SibSp","Parch","Fare","Embarked","Title")])
```
### Confusion Matrix
```{r}
confusionMatrix(pred_test , as.factor(y_test$Survived))
confusionMatrix(pred_train , as.factor(titanic.train$Survived))
```

### Cross validation
```{r}
tcontrol2 <- trainControl(method = "repeatedcv", number = 10, repeats = 10)
modelRF2 <- train(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title, ntree=100,
                 data=titanic.train, method = "rf",trControl = tcontrol2, tuneGrid=expand.grid(.mtry=c(10)))
modelRF2$results
```

```{r}
PassengerId <- titanic.test$PassengerId
output.df <- as.data.frame(PassengerId)
output.df$Survived <- pred_test

write.csv(output.df, file="kaggle_rf.csv", row.names = F)
```





